---
title: "Machine Learning II - MBD16 - Competition"
output:
  html_document: default
  html_notebook: default
---

#Table of content

1. Introduction and data preparation
2. Evaluation
3. Feature Engineering
4. Feature Selection
5. Model Selection
6. Evaluating Models

# 1. Introduction and data preparation

Read and map csv files to single data frame

```{r Setup, message = FALSE, eval=FALSE}

GitTestTestfinals

Test Mauri

df_labels <- read.csv("internal-training-labels.csv") # Read Labels file
df_values <- read.csv("internal-training-values.csv") # Read Values file
df <- merge(df_labels,df_values, by = "id") # Merge both files on 'ID'
rm(df_labels) # remove not used data frame
rm(df_values) # remove not used data frame

# first look at the data and it's structure
summary(df) 
str(df)

```

It can be observed, that there are various variables that miss a significant amount of fields. Here we handle this part.

```{r}

#population '0' values

#num_private '0' values

#latitude, longitude, height: check '0' values

```

# 2. Evaluation

Build Evaluation function that computes the accuracy rate of a model

Build CV function that splits the data set into k-folds for cross validation

```{r}



```

# 3. Feature Engineering

Build extra features
continuous to levels

https://github.com/drivendataorg/pump-it-up/blob/master/madRid/PumpItUp_DataDriven_2nd_place_madRid_team.R


```{r}

```

# 4. Feature Selection

Build feature selection part

```{r}

```

# 5. Model Selection

Build modeling or model selection part

```{r}

```

# 6. Model Evaluation

Store models and evaluations

```{r}

```

