---
title: "Machine Learning II - Group D - MBD16 - Competition"
output:
  html_document: default
  code_folding: show
  html_notebook: default
  pdf_document: default
---

#Instructions

Calling below function after running the RMD file will output a Data Frame with the 'id' variable and the prediction for 'status_group' including a header for the test dataset that lays on the filepath.

`evaluate('filepath')`


#Table of content

1. Data Exploration
2. Feature Engineering
3. Feature Selection
4. Model Selection

```{r message=FALSE, warning=FALSE, include=FALSE}

#install.packages("ggmap", type = "source")
library(data.table) # Data reading
library(mice) # imputing
library(plyr) # Data manipulation
library(dplyr) # Data manipulation
library(ggplot2) # Data visualization
library(caTools) # Split dataset
library(randomForest) # Modeling
library(caret) # Confusion matrix
library(maps) # Location visualization
library(mapdata) # Location visualization
library(ggmap) # Location visualization
library(geonames) # Get heights

```

```{r Setup, message=FALSE, warning=FALSE, include=FALSE}

wp_values <- fread("internal-training-values.csv",na.strings = '',stringsAsFactors = T) # Read Values file
wp_labels <- fread("internal-training-labels.csv",na.strings = '',stringsAsFactors = T) # Read Label file
waterPump <- tbl_df(merge(wp_values,wp_labels)) # Merge both files on 'ID'
rm(wp_labels) # remove not used data frame
rm(wp_values) # remove not used data frame


# Removing NAs from dataset
finalWP <- waterPump[complete.cases(waterPump),]
attach(finalWP)
attach(waterPump)

```

# 1. Data Exploration

Before starting engineering the features, the dataset is explored by plotting relevant features.

```{r warning=FALSE, echo=FALSE, comment=FALSE, message=FALSE}

# Distribution of classes to be predicted
table(status_group)
prop.table(table(status_group))

# Visualization of data
qplot(quantity, data=waterPump, geom="bar", fill=status_group) + 
  theme(legend.position = "right")
qplot(status_group, data=waterPump, geom="bar", fill=quantity) + 
  theme(legend.position = "right")

qplot(quality_group, data=waterPump, geom="bar", fill=status_group) + 
  theme(legend.position = "right")

qplot(waterpoint_type, data=waterPump, geom="bar", fill=status_group) + 
  theme(legend.position = "right") + 
  theme(axis.text.x=element_text(angle = -20, hjust = 0))

ggplot(subset(waterPump, construction_year > 0), aes(x =construction_year)) +
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)

# Visually mapping location data

Tanzania <- get_map("Tanzania", source = "google")
Tanzania <- get_map(location = c(longitude = 35, latitude = -7), zoom = 6)
bc = data.frame(waterPump$latitude, waterPump$longitude)
ggmap(Tanzania) + 
  geom_point(data = bc, mapping = aes(x = waterPump.longitude, y = waterPump.latitude, color = status_group)) + labs(x = 'longitude', y = 'latitude')

```

Above visualizations help choose relevant features that are possible strong predictors.

The value 'dry' of the 'quality' variable seems to isolate non functional waterpumps. Similarly, functional pumps tend to have a 'enough' value. This makes the 'quality' feature a possible strong predictor. 

# 2. Feature Engineering

An additional feature which represents the number of days since the data was obtained, is created.

Unrealistic, impossible and 'NA' values are imputed using mice and random forest.

Reducing factor levels for categorical variables 'installer' and 'funder'. Same values with different name variations are combined to a single level. Rare levels are combined into a single level 'Others'.


```{r, message=FALSE, warning=FALSE, comment=FALSE}

# Converting date to datetime class
finalWP$date_recorded <- as.Date(finalWP$date_recorded)

# Creating new feature from recorded date
finalWP$days_since_last_recorded <- max(finalWP$date_recorded)-finalWP$date_recorded
finalWP$days_since_last_recorded <- as.integer(finalWP$days_since_last_recorded)

# Selecting features with missing values for imputation
impWP <- select(finalWP,id,population,construction_year)
impWP$population[impWP$population == 0] <- NA
impWP$construction_year[impWP$construction_year == 0]<- NA

# Performing mice imputation, based on random forests
# miceMod <- mice(impWP[, !names(impWP) %in% "id"], method="rf")
```

```{r, message=FALSE, warning=FALSE, comment=FALSE, include=FALSE}
miceMod <- mice(impWP[, !names(impWP) %in% "id"], method="rf")
```

```{r, message=FALSE, warning=FALSE, comment=FALSE}
# Generating the completed data
miceOutput <- complete(miceMod)

# Imputing Height
options(geonamesUsername="mauricecream")

height <- finalWP %>% select(gps_height, longitude, latitude) %>% rowwise() %>% transmute(gps_height = ifelse(gps_height <= 0 , GNsrtm3(latitude,longitude)[,1], as.numeric(gps_height)))

finalWP$gps_height <- as.numeric(height$gps_height)


# Adding imputed values to dataset
finalWP <- select(finalWP,-c(population,construction_year))
finalWP <- cbind(finalWP,miceOutput)

# Converting funder & installer to lowercase
finalWP$funder <- as.character(finalWP$funder)
finalWP$installer <- as.character(finalWP$installer)
chr.cols <- finalWP %>% summarise_each(funs(is.character(.))) %>% unlist() %>% which() %>% names()
finalWP <- finalWP %>% mutate_each(funs(tolower), one_of(chr.cols))

# Installer - reducing factor levels
finalWP$installer <- as.factor(finalWP$installer)
finalWP$installer[finalWP$installer == "" | finalWP$installer == 0 | finalWP$installer == "-"] <- NA
finalWP$installer[finalWP$installer == "gove" | finalWP$installer == "gover" | finalWP$installer == "central government" | finalWP$installer == "central govt"] <- "government"
finalWP$installer[finalWP$installer == "commu"] <- "community"
finalWP$installer[finalWP$installer == "danid"] <- "danida"
finalWP$installer[finalWP$installer == "word" | finalWP$installer == "wo" | finalWP$installer == "word bank" | finalWP$installer == "world" | finalWP$installer == "wordl bank" | finalWP$installer == "would bank" | finalWP$installer == "world banks" | finalWP$installer == "world nk"] <- "world bank"

levels_installer = 11

installerNames <- names(summary(finalWP$installer)[1:levels_installer])
installer <- factor(finalWP$installer, levels=c(installerNames, "Other"))
installer[is.na(installer)] <- "Other"
finalWP$installer <- installer

# Funder - reduce factor levels
finalWP$funder <- as.factor(finalWP$funder)
finalWP$funder[finalWP$funder == "" | finalWP$funder == 0] <- NA

levels_funder = 16

funderNames <- names(summary(finalWP$funder)[1:levels_funder])
funder <- factor(finalWP$funder, levels=c(funderNames, "Other"))
funder[is.na(funder)] <- "Other"
finalWP$funder <- funder

# Imputing missing/incorrect latitudes & longitudes

finalWP$longitude[finalWP$lga =="Bariadi" & finalWP$longitude == 0] <- 34.33104
finalWP$latitude[finalWP$lga =="Bariadi" & finalWP$latitude == -0.00000002] <- -2.69166

finalWP$longitude[finalWP$lga =="Geita" & finalWP$longitude == 0] <- 32.23135
finalWP$latitude[finalWP$lga =="Geita" & finalWP$latitude == -0.00000002] <- -2.88504

finalWP$longitude[finalWP$lga =="Magu" & finalWP$longitude == 0] <- 33.25879
finalWP$latitude[finalWP$lga =="Magu" & finalWP$latitude == -0.00000002] <- -2.45705

# Setting missing values to False
finalWP$public_meeting <- ifelse(finalWP$public_meeting == "TRUE", "True", "False")
finalWP$public_meeting <- as.factor(finalWP$public_meeting)
finalWP$public_meeting[is.na(finalWP$public_meeting)] <- "False"

finalWP$permit <- ifelse(finalWP$permit == "TRUE", "True", "False")
finalWP$permit <- as.factor(finalWP$permit)
finalWP$permit[is.na(finalWP$permit)] <- "False"

# Age Feature
finalWP$age <- max(finalWP$construction_year)-finalWP$construction_year

# Date recorded features
finalWP$recorded_month <- finalWP$date_recorded

```

# 3. Feature Selection

Unneccesary, non-interpretable, constant or unclear features are removed. Feature importance and selection will be handled by random forest in the modeling section.

```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=FALSE}

# Removing redundant features
finalWP <- finalWP[, -which(names(finalWP) == "recorded_by")] # only one value (organization which recorded the information)
finalWP <- finalWP[, -which(names(finalWP) == "quantity_group")] # same as quantity
finalWP <- finalWP[, -which(names(finalWP) == "region_code")] # code for region
finalWP <- finalWP[, -which(names(finalWP) == "date_recorded")] # Date row was entered (not a factor determining functionality)
finalWP <- finalWP[, -which(names(finalWP) == "num_private")] #undefined,id field
finalWP <- finalWP[, -which(names(finalWP) == "district_code")]
finalWP <- finalWP[, -which(names(finalWP) == "quality_group")] #identical to water_quality
finalWP <- finalWP[, -which(names(finalWP) == "payment_type")] # similar to payment
finalWP <- finalWP[, -which(names(finalWP) == "scheme_management")] #similar to management
finalWP <- finalWP[, -which(names(finalWP) == "management_group")] #similar to management
finalWP <- finalWP[, -which(names(finalWP) == "source")] #similiar to source_type
finalWP <- finalWP[, -which(names(finalWP) == "subvillage")] #too many levels
finalWP <- finalWP[, -which(names(finalWP) == "wpt_name")] #too many levels
finalWP <- finalWP[, -which(names(finalWP) == "ward")]
finalWP <- finalWP[, -which(names(finalWP) == "lga")]
finalWP <- finalWP[, -which(names(finalWP) == "scheme_name")]
finalWP <- finalWP[, -which(names(finalWP) == "extraction_type")]
finalWP <- finalWP[, -which(names(finalWP) == "extraction_type_group")]
finalWP <- finalWP[, -which(names(finalWP) == "waterpoint_type_group")]
finalWP <- finalWP[, -which(names(finalWP) == "amount_tsh")]

```

# 4. Modeling Section

Using tuneRF function the optimal value 'mtry = 4' to use in a random forest algorithm is computed.

A model using Boosing trees algorithm gives a higher test error rate than Random forests.

```{r}

# train/test split
set.seed(1234)
sample <- sample.split(finalWP$status_group, SplitRatio = .7)
train <- subset(finalWP, sample == TRUE)
test <- subset(finalWP, sample == FALSE)

# Commented for notebook efficiency:
# Tuning Randomforest for optimal mtry parameter
#rf.all.tune <- tuneRF(finalWP[,-19], finalWP[,19], ntreeTry=800, stepFactor=1.5)

# Randomforest 2
set.seed(12345)
rf.all2 <- randomForest(status_group ~ .-id, mtry=4,ntree = 100,data=train,importance=TRUE)

varImpPlot(rf.all2)

yhat.rf2 <- predict(rf.all2 ,newdata=test,type = "response")

confusionMatrix(test$status_group,yhat.rf2)

```




```{r, include=FALSE}
evaluate <- function(filePath = file.choose()){
  DF  <- fread(filePath,na.strings = '',stringsAsFactors = T) # Read Values file
  evaluationtest <- tbl_df(DF)
  # Removing NAs from dataset
  attach(evaluationtest)
  
  # Converting date to datetime class
  evaluationtest$date_recorded <- as.Date(evaluationtest$date_recorded)
  
  # Creating new feature from recorded date
  evaluationtest$days_since_last_recorded <- max(evaluationtest$date_recorded)-evaluationtest$date_recorded
  evaluationtest$days_since_last_recorded <- as.integer(evaluationtest$days_since_last_recorded)
  
  # Selecting features with missing values for imputation
  imputation <- select(evaluationtest,id,population,construction_year)
  imputation$population[imputation$population == 0] <- NA
  imputation$construction_year[imputation$construction_year == 0]<- NA
  
  # Performing mice imputation, based on random forests
  miceMod <- mice(imputation[, !names(imputation) %in% "id"], method="rf")
  
  # Generating the completed data
  miceOutput <- complete(miceMod)
  
  # Adding imputed values to dataset
  evaluationtest <- select(evaluationtest,-c(population,construction_year))
  evaluationtest <- cbind(evaluationtest,miceOutput)
  
  # Converting funder & installer to lowercase
  evaluationtest$funder <- as.character(evaluationtest$funder)
  evaluationtest$installer <- as.character(evaluationtest$installer)
  chr.cols = evaluationtest %>% summarise_each(funs(is.character(.))) %>% unlist() %>% which() %>% names()
  evaluationtest = evaluationtest %>% mutate_each(funs(tolower), one_of(chr.cols))
  
  # Installer - reducing factor levels
  evaluationtest$installer <- as.factor(evaluationtest$installer)
  evaluationtest$installer[evaluationtest$installer == "" | evaluationtest$installer == 0 | evaluationtest$installer == "-"] <- NA
  evaluationtest$installer[evaluationtest$installer == "gove" | evaluationtest$installer == "gover" | evaluationtest$installer == "central government" | evaluationtest$installer == "central govt"] <- "government"
  evaluationtest$installer[evaluationtest$installer == "commu"] <- "community"
  evaluationtest$installer[evaluationtest$installer == "danid"] <- "danida"
  evaluationtest$installer[evaluationtest$installer == "word" | evaluationtest$installer == "wo" | evaluationtest$installer == "word bank" | evaluationtest$installer == "world" | evaluationtest$installer == "wordl bank" | evaluationtest$installer == "would bank" | evaluationtest$installer == "world banks" | evaluationtest$installer == "world nk"] <- "world bank"
  
  installerNames2 <- levels(finalWP$installer)
  installer <- factor(evaluationtest$installer, levels=c(installerNames2))
  installer[is.na(installer)] <- "Other"
  evaluationtest$installer <- installer
  
  # Funder - reduce factor levels
  evaluationtest$funder <- as.factor(evaluationtest$funder)
  evaluationtest$funder[evaluationtest$funder == "" | evaluationtest$funder == 0] <- NA
  
  funderNames2 <- levels(finalWP$funder)
  funder <- factor(evaluationtest$funder, levels=c(funderNames2))
  funder[is.na(funder)] <- "Other"
  evaluationtest$funder <- funder
  
  # Imputing missing/incorrect latitudes & longitudes
  evaluationtest$longitude[evaluationtest$lga =="Bariadi" & evaluationtest$longitude == 0] <- 34.33104
  evaluationtest$latitude[evaluationtest$lga =="Bariadi" & evaluationtest$latitude == -0.00000002] <- -2.69166
  
  evaluationtest$longitude[evaluationtest$lga =="Geita" & evaluationtest$longitude == 0] <- 32.23135
  evaluationtest$latitude[evaluationtest$lga =="Geita" & evaluationtest$latitude == -0.00000002] <- -2.88504
  
  evaluationtest$longitude[evaluationtest$lga =="Magu" & evaluationtest$longitude == 0] <- 33.25879
  evaluationtest$latitude[evaluationtest$lga =="Magu" & evaluationtest$latitude == -0.00000002] <- -2.45705
  
  # Setting missing values to False
  evaluationtest$public_meeting <- ifelse(evaluationtest$public_meeting == "TRUE", "True", "False")
  evaluationtest$public_meeting <- as.factor(evaluationtest$public_meeting)
  evaluationtest$public_meeting[is.na(evaluationtest$public_meeting)] <- "False"
  
  evaluationtest$permit <- ifelse(evaluationtest$permit == "TRUE", "True", "False")
  evaluationtest$permit <- as.factor(evaluationtest$permit)
  evaluationtest$permit[is.na(evaluationtest$permit)] <- "False"
  
  # Removing redundant features
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "recorded_by")] # only one value (organization which recorded the information)
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "quantity_group")] # same as quantity
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "region_code")] # code for region
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "date_recorded")] # Date row was entered (not a factor determining functionality)
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "num_private")] #undefined,id field
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "district_code")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "quality_group")] #identical to water_quality
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "payment_type")] # similar to payment
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "scheme_management")] #similar to management
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "management_group")] #similar to management
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "source")] #similiar to source_type
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "subvillage")] #too many levels
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "wpt_name")] #too many levels
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "ward")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "lga")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "scheme_name")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "extraction_type")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "extraction_type_group")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "waterpoint_type_group")]
  evaluationtest <- evaluationtest[, -which(names(evaluationtest) == "amount_tsh")]
  
  yhat.rffinal <- predict(rf.all2 ,newdata=evaluationtest,type = "response")
  submission <<- as.data.frame(yhat.rffinal)
  submission <<- cbind(DF$id, submission)
  names(submission) <- c('id','status_group')
  submission$status_group <- as.factor(submission$status_group)
  submission
}

```

